% =====================================================================
% STAGED INSERTS for SecureAI_Bond_2026.tex camera-ready.
% Numbers in <<...>> are filled from revision/out/harm_validation_analysis.json
% after the full NRP-panel run. Do NOT merge with placeholders unresolved.
% =====================================================================

% ---- INSERT 1: new subsection in Section III (after Harm Space and
%      Perturbation Categories), answering reviewer #1 (validate harm space). ----

\subsection{Harm-Space Validation}
\label{sec:validation}

A geometric framework is only as trustworthy as the coordinates it measures in.
We therefore validate that the seven harm dimensions are \emph{reliably
measurable} rather than asserted, via an inter-model agreement study over an
open, multi-family judge panel served by the NRP managed LLM API. We elicited
the full 7-D harm vector (each dimension $0$--$10$) from \textbf{<<NMODELS>>}
independent open models spanning <<NFAMILIES>> architecture families
(<<MODEL_LIST>>) on \textbf{<<NSCEN>>} scenarios, with three replications per
cell for test--retest estimation. Crucially this panel is disjoint from the
five systems under test and fully reproducible (no proprietary endpoints),
so dimension reliability is established independently of the models being
audited.

\begin{table}[htbp]
\caption{Harm-space reliability across the open judge panel ($k=$<<NMODELS>> models, $n=$<<NSCEN>> scenarios). ICC(2,$k$) is two-way random, absolute-agreement, average-measures; $\alpha$ is Krippendorff's interval alpha.}
\label{tab:validation}
\centering
\footnotesize
\begin{tabular}{@{}lcc@{}}
\toprule
\textbf{Harm dimension} & \textbf{ICC(2,$k$)} & \textbf{Krippendorff $\alpha$} \\
\midrule
Physical    & <<ICC_phys>>  & <<A_phys>>  \\
Emotional   & <<ICC_emot>>  & <<A_emot>>  \\
Financial   & <<ICC_fin>>   & <<A_fin>>   \\
Autonomy    & <<ICC_auto>>  & <<A_auto>>  \\
Trust       & <<ICC_trust>> & <<A_trust>> \\
Social      & <<ICC_soc>>   & <<A_soc>>   \\
Identity    & <<ICC_ident>> & <<A_ident>> \\
\midrule
\textbf{Overall} & \textbf{<<ICC_all>>} & \textbf{<<A_all>>} \\
\bottomrule
\end{tabular}
\end{table}

Aggregate inter-model reliability is <<ICC_all_word>> (ICC(2,$k$)$=$<<ICC_all>>,
Krippendorff $\alpha=$<<A_all>>), confirming that the seven dimensions index a
shared, model-independent structure rather than idiosyncratic per-model scales.
The dimensions cluster by reliability: <<HIGH_DIMS>> are measured most
consistently, while <<LOW_DIMS>> show the widest cross-model spread---a
direct empirical caveat for interpreting displacement on those axes.
Within-model test--retest reliability (median Pearson $r=$<<RETEST>> across
models) establishes the stochastic floor against which the perturbation effects
in Section~\ref{sec:results} are measured.

This validation also grounds the projection in Table~\ref{tab:deme}: the four
DEME v3 axes that map directly to harm dimensions (physical, autonomy,
legitimacy/trust, societal) are among the most reliably measured, supporting
the claim that the 7-D space preserves the essential structure of the
nine-dimensional moral vector.

% ---- INSERT 2: new subsection in Section VI (Discussion), answering
%      reviewer #1 (concrete end-to-end security case study). ----

\subsection{Worked Exploit: Salience-Driven Threshold Evasion}
\label{sec:exploit}

To make the deployment consequence concrete, we instantiate a content-moderation
filter that removes any item whose total harm score exceeds a threshold
$T$ on the $0$--$70$ scale, calibrated so that unperturbed judgments match the
audited gold labels. An attacker who controls only \emph{presentation}---not the
facts---applies euphemistic rewriting, a morally invariant transform by
construction. Across the panel, euphemistic rewriting lowers the total harm
score by a mean of <<EUPH_DELTA>> points (range <<EUPH_MIN>> to <<EUPH_MAX>>),
while dramatic rewriting raises it by <<DRAM_DELTA>> points.

The security consequence is a \emph{silent reclassification}: any item whose true
harm sits within the euphemistic drift band above $T$ is moved from
\textsc{remove} to \textsc{allow} without changing what actually happened. At the
calibrated threshold this flips \textbf{<<FLIP_PCT>>} of moderated decisions in
our gold set---an end-to-end exploit driven purely by salience manipulation. The
asymmetry compounds the risk: because the displacement is larger for euphemism
than for exaggeration, the filter is easier to \emph{evade} (hide real harm) than
to \emph{trigger} (manufacture false positives), the precise direction in which a
moderation system fails dangerously. Legal triage and clinical intake filters
that gate on a severity score inherit the same vulnerability.
